---
title: "Initial Korean covid 19"
author: "Anicia Friend"
date: "April 15, 2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  


```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  

The amount of females who got COVID and lived is more than males. However, out of the ones who got COVID and passed, the percentage of men who died is larger than the percent of woman who died. 


  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data = kRdata)
cat("Row Percentages:\n")
print(round(prop.table(table1, margin = 1) * 100, 1))
cat("\nColumn Percentages:\n")
print(round(prop.table(table1, margin = 2) * 100, 1))
```

Out of all the alive groups, 12% more women than men survived. Out of the dead group, about 30% more men than woman died. 98.7% of females lived, while 1.4% of the them died. 97% of men lived, while 3% of the men died. Percentage wise, COVID killed about twice the amount of men that it did women. 

## Inferential results 

```{r}
# Create the contingency table
table1 <- xtabs(~result + Sex, data = kRdata)

# Chi-Square Test
chi_result <- chisq.test(table1)

cat("Chi-Square Test Results:\n")
print(chi_result)

cat("\nExpected Counts:\n")
print(chi_result$expected)

cat("\nResiduals:\n")
print(chi_result$residuals)

# Fisher's Exact Test
fisher_result <- fisher.test(table1)

cat("\nFisher's Exact Test p-value:\n")
print(fisher_result$p.value)

```

In our Chi-Square result, we received a p-value of 0.0012. Since this is <0.05 (our point of significance), we reject our null (that dying is not independent of sex). In our Fishers test, we received a p-value of 0.0009 which is also <0.05, we again will be rejecting the null hypothesis. 

## Conclusions

The question we are addressing was if dying from COVID is independent of sex. We used a sample from South Korea and then used graphical, numerical, and inferential results to answer this question. From these results 98.7% of females lived, while 1.4% of them died and 97% of men lived, while 3% of the men died. Percentage wise, COVID killed about twice the amount of men than it did women. Eventually, we decided that our null hypothesis (that dying is not independent of sex) was rejected die to receiving p-values less than the value of significance(0.0012 and 0.0009). 


